---
title: "Email Compose"
author: "Katie Masiello"
date: "8/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(blastula)
library(ggplot2)
library(lubr)
library(palmerpenguins) # for fun data
```

Let's create some interesting data that we want to email.
```{r}
head(penguins)

# And let's plot something
penguin_plot <- ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Great data.  Now let's make an email with Blastula.

Simplest case:
```{r}
# note you can use **markdown** formatting with the `md()` function
message_1 <- blastula::compose_email(
  body = md("Hello, 
  
  That penguin project is looking swell.  You should see the data collected to date on our dashboard.
  
  Stay cool, 
  
  Katie")
)

message_1
```

Basic email sending:
Followed the instructions from https://rich-iannone.github.io/blastula/articles/sending_using_smtp.html
Since I don't want my email credentials being shared, I will pass those as an environment variable using the `pass_envvar` argument in `smtp_send`. See https://github.com/rich-iannone/blastula/issues/149 and PR https://github.com/rich-iannone/blastula/pull/159
```{r}
message_1 %>%
  smtp_send(
    from = "katie.masiello@rstudio.com",
    to = "katie.masiello@rstudio.com", #This will put everyone in the to line.  Use purrr::map() to iterate 
    subject = "Penguins",
    credentials = creds_envvar(
  user = Sys.getenv("SMTP_USER"),
  pass_envvar = "SMTP_PASSWORD",
  provider = "gmail")
  )

```


We can jazz it up with our plot:
```{r}
# Let's add some data
plot_email <- blastula::add_ggplot(penguin_plot) # this is a helper function to make plots images for emailing

message_2 <- blastula::compose_email(
  body = md(paste("Hello, <br>
                  Here is a plot of penguin data to date. <br>",
                  plot_email, 
                  "Stay cool, <br>
                  Katie")), 
  footer = md(paste0(
        "Email sent by using the `blastula` package on ", Sys.Date(), "."
      ))
  )

message_2
```

And send it with a little jazz in the subject line:
```{r}
days_remaining <- as.numeric(as.Date("2020-12-31")-Sys.Date()) #hypothetical timeline to end of penguin project

message_2 %>%
  smtp_send(
    from = "katie.masiello@rstudio.com",
    to = "katie.masiello@rstudio.com", #This will put everyone in the to line.  Use purrr::map() to iterate 
    subject = paste(days_remaining,
                    "days remain in the project and", 
                    nrow(penguins), "penguin measurements are complete"),
    credentials = creds_envvar(
  user = Sys.getenv("SMTP_USER"),
  pass_envvar = "SMTP_PASSWORD",
  provider = "gmail"
)
  )
```


And to get fancy, make it conditional:
```{r}
#Let's only send the email on weekdays
is_weekday <- (lubridate::wday(Sys.Date()) %in% c(2:6))

if(is_weekday){
  message_2 %>%
  smtp_send(
    from = "katie.masiello@rstudio.com",
    to = "katie.masiello@rstudio.com", #This will put everyone in the to line.  Use purrr::map() to iterate 
    subject = paste(days_remaining,
                    "days remain in the project and", 
                    nrow(penguins), "penguin measurements are complete"),
    credentials = creds_envvar(
  user = Sys.getenv("SMTP_USER"),
  pass_envvar = "SMTP_PASSWORD",
  provider = "gmail"
)
  )
}
```

We should automate this:
see https://github.com/kmasiello/email_distribution